A Comparision of Bootstrap Methods

### 2022-11-30 project/PaperTemplate/paper.Rmd
Added the examples

```{r setup, include = FALSE}
library(tidyverse)
library(mosaic)
library(mdsr)
library(knitr)
library(kableExtra)
```

@@

@@

@@

@@

@@

@@

>>>>

```{r setup, include = FALSE}
library(tidyverse)
library(mosaic)
library(mdsr)
library(knitr)
library(kableExtra)
library(boot)
```

@@

We demonstrate performing a non-parametric bootstrap below using `SnowGR` which gives the official snowfall dataset by month for Grand Rapids, MI, starting in 1893. We will be using the `Dec` variable which gives the number of inches of snow that fell in December of each year. In particular, we are interested in the sample mean and later finding confidence intervals for population mean of `Dec`. The distribution is below.

```{r}
data(SnowGR)
favstats(~ Dec, data = SnowGR)
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", title = "Distribution of Annual Snowfall in Decemeber", y = "Count")
```

Next, we will perform the non-parametric bootstrap using 10,000 replications.

orig_mean <- mean(~ Dec, data = SnowGR) # theta hat

set.seed(495)
nboot <- 10000

mean <- rep(0,nboot)
sd <- rep(0,nboot)
se <- rep(0,nboot)
t_stat <- rep(0,nboot)

for (i in 1:nboot) {
  resampled <- as.data.frame(mosaic::resample(SnowGR, replace = TRUE))
  mean[i] <- mean(~Dec, data = resampled)
  sd[i] <- sd(~Dec, data = resampled)
 # se[i] <- sd[i]/nrow(resampled)
#  t_stat[i] <- (mean[i] - orig_mean)/se[i]
}

dec_means <- as.data.frame(mean)
#dec_t_stat <- as.data.frame(t_stat)

#set.seed(495)
#dec_means <- do(10000)*mean(~ Dec, data = mosaic::resample(SnowGR, replace = TRUE))

gf_dhistogram(~ mean, data = dec_means) %>%
  gf_dens() %>%
  gf_labs(title = "Histogram of 10,000 Bootstrapped Mean Dec Values")
```

@@

Using our example, to find a 95% confidence interval for the population mean of `Dec`, we would get the following:

```{r}
mean(~ Dec, data = SnowGR) + qnorm(c(0.025, 0.975)) * sd(~ Dec, data = SnowGR) / sqrt(nrow(SnowGR))
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.85639)) +
  geom_vline(aes(xintercept= 17.65957))
```

@@

Finding a 95% confidence interval for the true population mean of `Dec` using the percentile method, we get:

```{r}
qdata(~ mean, c(0.025, 0.975), data = dec_means)
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.91170)) +
  geom_vline(aes(xintercept= 17.72353))
```
This interval is similar to the one created using the Standard method, but shifted slightly to the right.

@@

Using this method to create a 95% confidence interval has a couple more steps because we need to find the bias-correction value. First we calculate the sample mean.

```{r}
sample_mean <- mean(~Dec, data = SnowGR); sample_mean
```

Then we find the proportion of bootstrap replications that have a sample mean less than the original sample mean.
```{r}
less_than_sample_mean <- sum(ifelse(dec_means <= sample_mean, 1, 0))/10000
less_than_sample_mean
```

From this, we can calculate the bias-correction value, $z_0$:
```{r}
z0 <- qnorm(less_than_sample_mean); z0
```

Then we can find the modified percentiles:

```{r}
alphalow <- 0.025; alphahigh <- 0.975

newlow <- pnorm(2*z0 + qnorm(alphalow)); #newlow
newhigh <- pnorm(2*z0 + qnorm(alphahigh)); #newhigh

qdata(~ mean, c(newlow, newhigh), data = dec_means)
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 14.01172)) +
  geom_vline(aes(xintercept= 17.82017))

@@

In practice, we can use the jackknife procedure to estimate the acceleration value for this method. Thus, we will run the jackknife procedure:

```{r}
theta <- function(x){mean(x)}
jackknife_results <- bootstrap::jackknife(SnowGR$Dec, theta) 
```

We find the mean of the jackknife values below:

```{r}
jackmean <- mean(~ jackknife_results$jack.values); jackmean
```

Then we can compute $a$.

```{r}
estimated_a <- (1/6)*sum((jackknife_results$jack.values - jackmean)^3)/
  (sum((jackknife_results$jack.values - jackmean)^2))^(1.5); estimated_a
```

Now, we can find the adjust percentiles and the bias-corrected with acceleration confidence interval.

```{r}
a_newlow <- pnorm(z0 + (z0 + qnorm(alphalow))/(1-estimated_a*(z0 + qnorm(alphalow)))); #a_newlow
a_newhigh <- pnorm(z0 + (z0 + qnorm(alphahigh))/(1-estimated_a*(z0 + qnorm(alphahigh)))); #a_newhigh
qdata(~ mean, c(a_newlow, a_newhigh), data = dec_means)
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.91591)) +
  geom_vline(aes(xintercept= 17.72353))

@@

We can use the following to get our studentized confidence interval:
```{r}
mean(~ Dec, data = SnowGR) + qt(c(0.025, 0.975), df = (nrow(SnowGR-1))) * 
  sd(~ Dec, data = SnowGR) / sqrt(nrow(SnowGR))
```

orig_mean <- mean(~ Dec, data = SnowGR) # theta hat
orig_se <- sd(~ Dec, data = SnowGR)/sqrt(nrow(SnowGR))

nboot <- 10000

se <- rep(0,nboot)
t_stat <- rep(0,nboot)

for (i in 1:nboot) {
  se[i] <- sd[i]/sqrt(nrow(SnowGR))
  t_stat[i] <- (mean[i] - orig_mean)/se[i]
}

dec_t_stat <- as.data.frame(t_stat)

q <- unname(quantile(dec_t_stat$t_stat, c(0.025, 0.975)))
lower <- q[1]
upper <- q[2]

c(orig_mean - upper*orig_se, orig_mean - lower*orig_se)
